---
title             : "Final Project: Milestone 3"
shorttitle        : "Milestone 3"
author: 
  - name          : "Gerardo Palacios"
    affiliation   : "1,2"
affiliation:
  - id            : "1"
    institution   : "DSC 425 - Time Series Analysis and Forecasting"
  - id            : "2"
    institution   : "DePaul University"
floatsintext      : yes
figsintext        : yes
figurelist        : yes
tablelist         : yes
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            :  papaja::apa6_pdf
fontsize          : 11pt
header-includes:
  - \usepackage{setspace}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage[backend=biber]{biblatex}
  - \addbibresource{r-references.bib}
  - \usepackage{afterpage}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
editor_options: 
  chunk_output_type: console
---

\singlespacing

```{r setup, include = FALSE,echo=FALSE}
library("papaja")
library(lubridate)
library(tidyquant)
library(zoo)
library(knitr)
library(kableExtra)
library(xtable)
library(ggfortify)
```

# Crypto-Currency and Stock Data - Ethereum and NVIDIA

As a group we are looking at a set of 7 stocks and crypto-currencies, with the focus of finding any relationships between them. My individual contribution will be to look at a stock and crypto-currency set, Ethereum and NVIDIA. NVIDIA is a technology company that manufacture high-end GPU's. A majority of their revenues are spearheaded through GPU card sales which are subsequently used for crypto-mining. This suggests there may be a relationship between the two series. 


The following data is retrieved using a Yahoo API in R named tidyquant. Retrieving the daily adjusted stock prices for NVIDIA and Ethereum between 10-01-2016 to 09-30-2021. The biggest difference between the two series is the number of instances. Since NvIDIA is a stock, it is subject to the market closing on weekends, holidays, and business hours, whereas a crypto-currency is actively traded 24/7. This means that there will be more instances for Ethereum than NVIDIA within the same time-frame.



```{r echo=FALSE, message=FALSE, warning=FALSE}
start_date = '2016-10-01'
end_date = '2021-09-30'
tickers = c('SPY','AMD','NVDA','BTC-USD','ETH-USD','LINK-USD','COIN','ADA-USD')

# Read csv file if exists
if(file.exists('tickers.csv')){
  stocks = read.csv('tickers.csv')
  stocks$date = ymd(stocks$date)
} else{ # Download and write if csv does not exist
  stocks = tq_get(tickers,
                  from=start_date,
                  to=end_date
                  )[,c('date','symbol','volume','adjusted')]
  stocks$date = ymd(stocks$date)
  write.csv(stocks,'tickers.csv', row.names = F)
}

# A function that will return only one stock as zoo object
get_stock_zoo <- function(ticker){
    selected_stock = stocks[stocks$symbol==ticker,]
    adjusted.price = zoo(
      selected_stock$adjusted,
      order.by = selected_stock$date
      )
    return(adjusted.price)
}

get_stock_df <- function(ticker){
    selected_stock = stocks[stocks$symbol==ticker,]
    return(selected_stock)
}

```

The NVIDA series contains 1,257 observations and 4 features, date, symbol, volume and adjusted price. 
Ethereum contains 1,826 observations with the same four features (date, symbol, volume, adj. price).

```{r exp1, include=T,echo=F,warning=FALSE}
df_nvd <- get_stock_df('NVDA')
df_eth <- get_stock_df('ETH-USD')
```
 


## Distributions
The distributions of the adjusted prices for two series shown in Figure 1 are highly skewed right, more often the prices are lower than extremely high. Aside from the skewness, NVIDA has 4 different peaks in its distribution suggesting multiple means during different time periods compared to Ethereum which only has a single peak.

```{r dis1, include=T,echo=F,warning=FALSE,fig.height=1.75,fig.fullwidth=T,fig.align="center",fig.pos='htb!'}

eth <- ggplot(data=df_eth,aes(x=adjusted)) + 
  geom_histogram() + ggtitle('Ethereum\nDistribution of Adjusted Price')

nvd <- ggplot(data=df_nvd,aes(x=adjusted)) + 
  geom_histogram() + ggtitle('NVIDIA\nDistribution of Adjusted Price')

cowplot::plot_grid(
  eth,nvd
)

```

## Graphing the Time Series
```{r data1, include=T,echo=F,warning=FALSE}
df_eth <- get_stock_zoo('ETH-USD')
df_nvd <- get_stock_zoo('NVDA')
```


Figure 2 shows the time series for Ethereum and NVIDIA over the past 5 years. It appears to be a multiplicative, non-stationary time series with an exponential positive trend that has exploded most recently in 2021. What is also apparent is that both series seem to follow a similar trend. There is a are two similar peaks that happen in the same time frame.

```{r data2, include=T,echo=F,warning=FALSE,fig.height=1.75,fig.fullwidth=T,fig.align="center",fig.pos='htb!'}
cowplot::plot_grid(
 autoplot(df_eth)+ggtitle('Ethereum Price Timeseries'),autoplot(df_nvd)+ggtitle('NVIDIA Price TimeSeries') 
)

```
\newpage
Figure 3, we can see the time series with a log transformation. It has transformed the exponential behavior into something more linear. There still remains a general increasing trend, and appears to be more additive. Again, we can note that both series are very similar trends.

```{r data3, include=T,echo=F,warning=FALSE,fig.height=1.75,fig.fullwidth=T,fig.align="center",fig.pos='htb!'}
cowplot::plot_grid(
 autoplot(log(df_eth))+ggtitle('Ethereum Price Timeseries'),autoplot(log(df_nvd))+ggtitle('NVIDIA Price TimeSeries') 
)

```

Figure 4, we can see the log returns. The plot shows general white noise with a few outliers in 2017 and 2020.They both seem to have a mean that floats around zero, and have a similar range between .2 and -.2 (with the exception of an outlier in 2020 at -.4)


```{r data4, echo=F, fig.align="center", fig.fullwidth=T, fig.height=1.75, warning=FALSE,fig.pos='htb!'}

cowplot::plot_grid(
 autoplot(diff(log(df_eth)))+ggtitle('Ethereum'),
 autoplot(diff(log(df_nvd)))+ggtitle('NVIDIA') 
)

```

# Auto-correlation

## ACF of the time series

Figure 4 is the ACF plot for both Ethereum and NVIDIA. Auto-correlation has a strong presence in this time series. The ACF gradually decreases indicating a non-stationary series.

```{r data5a, include=T,echo=F,warning=F,fig.height=1.75,fig.fullwidth=T,fig.align="center",fig.pos='htb!'}
conf.level <- 0.95
ciline_eth <- qnorm((1 - conf.level)/2)/sqrt(length(df_eth))
ciline_nvd <- qnorm((1 - conf.level)/2)/sqrt(length(df_nvd))

nvd_acf <- acf((log(df_nvd)), na.action = na.pass,plot = FALSE)
nvd_acf <- with(nvd_acf, data.frame(lag, acf))


eth_acf <- acf((log(df_eth)), na.action = na.pass,plot = FALSE)
eth_acf <- with(eth_acf, data.frame(lag, acf))

eth <- ggplot(data = eth_acf, mapping = aes(x = lag, y = acf)) +
       geom_hline(aes(yintercept = 0)) +
       geom_segment(mapping = aes(xend = lag, yend = 0)) + 
       ggtitle("Ethereum Log(Price) ACF") + 
       geom_hline(aes(yintercept = ciline_eth),linetype = 3, color = 'darkblue') + 
       geom_hline(aes(yintercept = -ciline_eth), linetype = 3, color = 'darkblue')

nvd <- ggplot(data = nvd_acf, mapping = aes(x = lag, y = acf)) +
       geom_hline(aes(yintercept = 0)) +
       geom_segment(mapping = aes(xend = lag, yend = 0)) + ggtitle("NVIDIA Log(Price) ACF")+ 
       geom_hline(aes(yintercept = ciline_nvd),linetype = 3, color = 'darkblue') + 
       geom_hline(aes(yintercept = -ciline_nvd), linetype = 3, color = 'darkblue')

cowplot::plot_grid(
  eth, nvd
)

```

## PACF of the time series
Figure 5 shows the PACF for both Ethereum and NVIDIA. Ethereum shows auto-correlation with an extreme value at lag 19, while NVIDIA shows no evidence of partial auto-correlation.

```{r data5b, include=T,echo=F,warning=F,fig.height=1.75,fig.fullwidth=T,fig.align="center",fig.pos='htb!'}

nvd_pacf <- pacf(log(df_nvd), na.action = na.pass,plot=FALSE)

nvd_pacf <- with(nvd_pacf, data.frame(lag, acf))


eth_pacf <- pacf((log(df_eth)), na.action = na.pass,plot=FALSE)
eth_pacf <- with(eth_pacf, data.frame(lag, acf))

eth <- ggplot(data = eth_pacf, mapping = aes(x = lag, y = acf)) +
       geom_hline(aes(yintercept = 0)) +
       geom_segment(mapping = aes(xend = lag, yend = 0)) + 
       ggtitle("Ethereum Log(Price) PACF") + 
       geom_hline(aes(yintercept = ciline_eth),linetype = 3, color = 'darkblue') + 
       geom_hline(aes(yintercept = -ciline_eth), linetype = 3, color = 'darkblue')


nvd <- ggplot(data = nvd_pacf, mapping = aes(x = lag, y = acf)) +
       geom_hline(aes(yintercept = 0)) +
       geom_segment(mapping = aes(xend = lag, yend = 0)) + 
       ggtitle("NVIDIA Log(Price) PACF") + 
       geom_hline(aes(yintercept = ciline_eth),linetype = 3, color = 'darkblue') + 
       geom_hline(aes(yintercept = -ciline_eth), linetype = 3, color = 'darkblue')


cowplot::plot_grid(
  eth, nvd
)
```


## ACF of the Log Returns
Figure 5 shows the ACF for the log returns of both Ethereum and NVIDIA. Ethereum shows some minor auto-correlation, while NVIDIA shows evidence of seasonality and auto-correlation.

```{r data5c, include=T,echo=F,warning=F,fig.height=1.75,fig.fullwidth=T,fig.align="center",fig.pos='htb!'}
conf.level <- 0.95
ciline_eth <- qnorm((1 - conf.level)/2)/sqrt(length(df_eth))
ciline_nvd <- qnorm((1 - conf.level)/2)/sqrt(length(df_nvd))

nvd_acf <- acf((diff(log(df_nvd))), na.action = na.pass,plot = FALSE)
nvd_acf <- with(nvd_acf, data.frame(lag, acf))


eth_acf <- acf((diff(log(df_eth))), na.action = na.pass,plot = FALSE)
eth_acf <- with(eth_acf, data.frame(lag, acf))

eth <- ggplot(data = eth_acf, mapping = aes(x = lag, y = acf)) +
       geom_hline(aes(yintercept = 0)) +
       geom_segment(mapping = aes(xend = lag, yend = 0)) + 
       ggtitle("Ethereum diff(Log(Price)) ACF") + 
       geom_hline(aes(yintercept = ciline_eth),linetype = 3, color = 'darkblue') + 
       geom_hline(aes(yintercept = -ciline_eth), linetype = 3, color = 'darkblue')

nvd <- ggplot(data = nvd_acf, mapping = aes(x = lag, y = acf)) +
       geom_hline(aes(yintercept = 0)) +
       geom_segment(mapping = aes(xend = lag, yend = 0)) + ggtitle("NVIDIA diff(Log(Price)) ACF")+ 
       geom_hline(aes(yintercept = ciline_nvd),linetype = 3, color = 'darkblue') + 
       geom_hline(aes(yintercept = -ciline_nvd), linetype = 3, color = 'darkblue')

cowplot::plot_grid(
  eth, nvd
)

```

\newpage
## Ljung Box Test

This can be further confirmed by performing the Ljung Box test. At lag 1, the Ljung-Box p-value is close to zero. This indicates that at the 99% confidence, the null hypothesis is rejected and one can conclude that the series is not independently distributed and exhibit serial correlation.

```{r data7, include=T,echo=F,warning=F}

Box.test(log(df_nvd), lag = 1, type = "Ljung-Box")
Box.test(log(df_eth), lag = 1, type = "Ljung-Box")

```

